---
title: "Creating ADRS with Lugano 2014 Criteria"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating ADRS with Lugano 2014 Criteria}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(admiraldev)
library(gt)
```

# Introduction

This article describes creating an `ADRS` ADaM dataset for lymphoma studies based on the [**Lugano 2014 response criteria**](https://doi.org/10.1200/JCO.2013.54.8800).

Lymphoma response assessment under Lugano 2014 is based on a combination of imaging-based evaluations:

-   **PET-CT based assessment**, providing metabolic response evaluation using the 5-point Deauville scale.
-   **CT-based assessment**, providing anatomic response evaluation of nodal and extranodal disease.

Depending on the study and the disease subtype, response evaluation may use PET-CT as the primary modality (for FDG-avid lymphomas) or CT alone (for non–FDG-avid lymphomas). Some studies collect both PET-CT and CT response components, and a combined overall response is derived.

Please check [Lugano 2014 Classification](https://imagingendpoints.com/wp-content/uploads/2022/07/IEP-6936-Lugano-at-IE-2022-FV1.0_DIGITAL_Final-25-07-22.pdf) for more details.

Note that only the Lugano 2014-specific derivation of the combined overall timepoint response is covered in this vignette, specifically for scenarios where the combined overall response is not collected directly on the CRF. For extended guidance on common steps in ADRS creation and additional response endpoints, refer to the examples in [Creating ADRS (Including Non-standard Endpoints)](adrs.html).

# Lugano 2014 Response Categories for Lymphoma

The Lugano 2014 response criteria define lymphoma response using PET-CT based metabolic assessment and CT-based anatomic assessment.

The following tables summarize the response categories used in this vignette for PET-CT and CT assessments. These summaries are intended to support the example derivations and should be aligned with the study protocol and statistical analysis plan.

## PET-CT Based Response Categories

```{r pet-response-categories, echo=FALSE, message=FALSE}
pet_response_categories <- tibble::tribble(
  ~`PET-CT Response`, ~`Description`, ~`Derived Overall Category`,
  "CMR", "Complete metabolic response", "CR",
  "PMR", "Partial metabolic response", "PR",
  "SMD", "Stable metabolic disease", "SD",
  "NMR", "No metabolic response", "SD",
  "PMD", "Progressive metabolic disease", "PD",
  "NE",  "Not evaluable", "NE or fallback logic",
  "ND",  "Not done or not determined", "Fallback to CT if available"
)

pet_response_categories |>
  gt::gt() |>
  gt::tab_header(
    title = "Table 1: PET-CT Based Response Categories",
    subtitle = "Lugano 2014 response categories used in this vignette"
  )
```

## CT-Based Response Categories

```{r ct-response-categories, echo=FALSE, message=FALSE}
ct_response_categories <- tibble::tribble(
  ~`CT Response`, ~`Description`, ~`Derived Overall Category`,
  "CR", "Complete response", "CR",
  "PR", "Partial response", "PR",
  "SD", "Stable disease", "SD",
  "PD", "Progressive disease", "PD",
  "NE", "Not evaluable", "NE",
  "ND", "Not done or not determined", "ND",
  "NED", "No evidence of disease", "NED"
)

ct_response_categories |>
  gt::gt() |>
  gt::tab_header(
    title = "Table 2: CT-Based Response Categories",
    subtitle = "Anatomic response categories used in this vignette"
  )
```

In the example data, `SMD` is used for stable metabolic disease. Some implementations may use `NMR` for no metabolic response. For the purpose of the combined overall response derivation in this vignette, both `SMD` and `NMR` are mapped to `SD`.

Values such as `NED`, `PSP`, and `ND` may require study-specific handling. The simplified derivation implemented in this vignette focuses on deriving a combined overall response from available PET-CT and CT response records.

# Programming Workflow

-   [Read in Data](#readdata)
-   [Pre-processing of Input Records](#input)
-   [Derive PET-CT and CT Response Parameters](#param)
-   [Derive Combined Overall Timepoint Response (OVRLRESC)](#ovrlresc)
-   [Other Endpoints](#other)

## Required Packages

The examples of this vignette require the following packages.

```{r, warning=FALSE, message=FALSE}
library(admiral)
library(admiralonco)
library(pharmaversesdtm)
library(pharmaverseadam)
library(dplyr)
library(tibble)
```

## Read in Data {#readdata}

To begin, all data frames needed for the creation of `ADRS` should be read into the environment. This will be a company-specific process. For this vignette, the main input datasets are `ADSL` and `RS`.

For demonstration purposes, the SDTM and ADaM datasets based on CDISC Pilot test data from `{pharmaversesdtm}` and `{pharmaverseadam}` are used.

In this vignette, the `RS` SDTM dataset is expected to contain lymphoma response assessments based on Lugano 2014 criteria. The example `RS` dataset contains separate records for:

-   PET-CT based response assessments, identified by `RSSCAT = "INCLUDING PET-CT SCAN"`.
-   CT-based response assessments, identified by `RSSCAT = "NOT INCLUDING PET SCAN"`.

```{r message=FALSE}

# Lymphoma SDTM data
#rs <- pharmaversesdtm::rs_onco_lymphoma
rs <- rs_onco_lymphoma

# Convert blanks to NA
rs <- convert_blanks_to_na(rs)

# ADaM data
adsl <- pharmaverseadam::adsl
```

```{r echo=FALSE}
# select subjects from adsl such that there is one subject without RS data
rs_subjects <- unique(rs$USUBJID)
adsl_subjects <- unique(adsl$USUBJID)
adsl <- filter(
  adsl,
  USUBJID %in% union(rs_subjects, setdiff(adsl_subjects, rs_subjects)[1])
)
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
   rs,
  display_vars = exprs(USUBJID, RSTESTCD,RSCAT,RSSCAT,RSMETHOD, RSSTRESC, VISIT, VISITNUM, RSDTC)
)
```

## Pre-processing of Input Records {#input}

At this step, it may be useful to join `ADSL` to your `RS` domain. Only the `ADSL` variables used for derivations are selected at this step.

```{r eval=TRUE}
adsl_vars <- exprs(TRTSDT)
adrs <- derive_vars_merged(
  rs,
  dataset_add = adsl,
  new_vars = adsl_vars,
  by_vars = get_admiral_option("subject_keys")
)

```

### Partial Date Imputation and Deriving `ADT`, `ADTF`, `AVISIT`, `AVISITN` etc.

If your data collection allows for partial dates, you could apply a company-specific imputation rule at this stage when deriving `ADT`. For this example, here we impute missing day to last possible date.

```{r}
adrs <- adrs %>%
  derive_vars_dtm(
    dtc = RSDTC,
    new_vars_prefix = "A",
    highest_imputation = "D",
    date_imputation = "last"
  ) %>%
  derive_vars_dtm_to_dt(exprs(ADTM)) %>%
  derive_vars_dy(
    reference_date = TRTSDT,
    source_vars = exprs(ADT)
  ) %>%
  mutate(
    AVISIT = VISIT,
    AVISITN = VISITNUM
  )

```

### Derive `PARAMCD`, `PARAM`, `PARAMN`

In this `RS` dataset, both PET-CT and CT response records use `RSTESTCD = "OVRLRESP"` and are distinguished by `RSSCAT` and `RSMETHOD`. For this vignette, `RSTESTCD` and `RSSCAT` are used to derive `PARAMCD`, `PARAM`, and `PARAMN`.

```{r, eval=TRUE, include=TRUE, message=FALSE}
# Prepare param_lookup for SDTM RSTESTCD and RSSCAT to add metadata
param_lookup <- tibble::tribble(
  ~RSTESTCD,  ~RSSCAT,                   ~PARAMCD, ~PARAM,          ~PARAMN,
  "OVRLRESP", "INCLUDING PET-CT SCAN",   "PETRSP", "PET-CT Response", 1,
  "OVRLRESP", "NOT INCLUDING PET SCAN",  "CTRSP",  "CT Response",     2
)

adrs <- adrs %>%
  derive_vars_merged_lookup(
    dataset_add = param_lookup,
    by_vars = exprs(RSTESTCD, RSSCAT)
  ) %>%
  mutate(
    PARCAT1 = RSCAT,
    AVALC = case_when(
      RSSTAT == "NOT DONE" ~ "ND",
      TRUE ~ RSSTRESC
    )
  )
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
   adrs,
  display_vars = exprs(USUBJID,PARAMCD,PARAM,AVALC, AVISIT, AVISITN,ADT)
)
```

## Derive Combined Overall Timepoint Response (`OVRLRESC`) Parameter {#ovrlresc}

For this vignette, the combined overall timepoint response parameter, `OVRLRESC`, is derived from the PET-CT and CT response records collected at each visit.

This example represents a scenario where the combined overall response is `not collected directly` on the CRF. Instead, it is derived using the available PET-CT and CT response records.

### General Derivation Assumptions Used in This Vignette

The following table summarizes the assumptions used in this vignette to derive the combined overall timepoint response from PET-CT and CT response records under Lugano 2014. These assumptions are intended for demonstration purposes. Please refer to your study protocol, statistical analysis plan, and other study documentation before using in production analyses.

#### Table: Combined Overall Timepoint Response Based on Lugano 2014 Response Categories

```{r , eval=TRUE, include=TRUE, message=FALSE, echo=FALSE}
combined_lugano_table <- tibble::tribble(
  ~`PET-CT Response`,                         ~`CT Response`,            ~`Combined Overall Response`,
  "CMR",                                      "Any",                     "CR",
  "PMR",                                      "Any",                     "PR",
  "SMD or NMR",                               "Any non-progressive",     "SD",
  "PMD",                                      "Any",                     "PD",
  "NE",                                       "NE",                      "NE",
  "NE, with prior evaluable PET-CT",          "Non-progressive",         "Carry forward prior PET-CT response",
  "NE, no prior evaluable PET-CT",            "CR / PR / SD / NED",      "Use current CT response",
  "NE",                                       "PD",                      "PD",
  "NED",                                      "CR / PR / SD / PD / NE",  "NED",
  "NE",                                       "NED",                     "NED",
  "NED",                                      "NED",                     "NED",
  "ND or not done",                           "CR / PR / SD / PD / NED", "Use current CT response",
  "PSP",                                      "Any",                     "PSP",
  "Missing or ND",                            "Missing or NE",           "ND"
)

combined_lugano_table |>
  gt::gt() |>
  gt::tab_header(
    title = "Table 3: Combined Overall Timepoint Response",
    subtitle = "PET-CT, CT, and Combined Overall Response Mapping"
  ) |>
  gt::tab_source_note(
    source_note = "Prior PET-CT carry-forward applies only when the current CT response is non-progressive."
  ) |>
  gt::tab_source_note(
    source_note = "When PET-CT is NE and no prior evaluable PET-CT exists, the current CT response is used if available."
  ) |>
  gt::tab_source_note(
    source_note = "When PET-CT is NE and CT shows PD, study-specific confirmation may apply, such as confirmatory imaging or biopsy. This is not implemented in this example."
  ) |>
  gt::tab_source_note(
    source_note = "Pseudoprogression handling, including final-visit conversion, is study-specific and not implemented in this example."
  )
```

##### Combined Overall Time Point Response(`OVRLRESC`) Records referenced from above table.

###### Derive prior evaluable PET-CT response for carry-forward logic

```{r, eval=TRUE, message=FALSE, include=TRUE}

adrs_prior_pet <- adrs %>%
  filter(PARAMCD == "PETRSP") %>%
  derive_vars_joined(
    dataset_add = adrs %>%
      filter(PARAMCD == "PETRSP",
             AVALC %in% c("CMR", "PMR", "SMD", "NMR", "PMD", "NED", "PSP")) %>%
      select(!!!get_admiral_option("subject_keys"),ADT,AVISITN,AVALC) %>%
      distinct(),
    by_vars = get_admiral_option("subject_keys"),
    order = exprs(ADT, AVISITN),
    mode = "last",
    join_type = "before",
    filter_join = ADT.join < ADT,
    new_vars = exprs(
     AVALC_P = AVALC,
     ADT_P = ADT
    )
  )

adrs <- bind_rows(
  adrs_prior_pet,
  adrs %>% filter(PARAMCD != "PETRSP")
)
```

###### Derive Combined Overall Time Point Response.

```{r eval=TRUE, message=FALSE, include=TRUE}
adrs <- derive_param_computed(
  dataset = adrs,
  by_vars = exprs(!!!get_admiral_option("subject_keys"), !!!adsl_vars, DOMAIN, ADT, ADY, ADTM, ADTF, VISIT, VISITNUM, AVISIT, AVISITN),
  parameters = c("PETRSP", "CTRSP"),
  set_values_to = exprs(
    AVALC = case_when(
      # PET-CT evaluable: PET-CT metabolic response determines overall response
      AVALC.PETRSP == "CMR" ~ "CR",
      AVALC.PETRSP == "PMR" ~ "PR",
      AVALC.PETRSP %in% c("SMD", "NMR") ~ "SD",
      AVALC.PETRSP == "PMD" ~ "PD",

      # PET-CT special responses
      AVALC.PETRSP == "NED" ~ "NED",
      AVALC.PETRSP == "PSP" ~ "PSP",

      # PET-CT is NE and CT indicates progression or special status
      AVALC.PETRSP == "NE" & AVALC.CTRSP == "PD" ~ "PD",
      AVALC.PETRSP == "NE" & AVALC.CTRSP == "NED" ~ "NED",
      AVALC.PETRSP == "NE" & AVALC.CTRSP == "NE" ~ "NE",

      # PET-CT is NE and prior evaluable PET-CT exists;
      # carry forward prior PET-CT response if current CT is non-progressive
      AVALC.PETRSP == "NE" & !is.na(AVALC_P.PETRSP) &
        (is.na(AVALC.CTRSP) | AVALC.CTRSP %in% c("CR", "PR", "SD", "NE", "NED")) ~ case_when(
          AVALC_P.PETRSP == "CMR" ~ "CR",
          AVALC_P.PETRSP == "PMR" ~ "PR",
          AVALC_P.PETRSP %in% c("SMD", "NMR") ~ "SD",
          AVALC_P.PETRSP == "PMD" ~ "PD",
          AVALC_P.PETRSP == "NED" ~ "NED",
          AVALC_P.PETRSP == "PSP" ~ "PSP",
          TRUE ~ NA_character_
        ),

      # PET-CT is NE and no prior evaluable PET-CT exists; use current CT response
      AVALC.PETRSP == "NE" & is.na(AVALC_P.PETRSP) &
        AVALC.CTRSP %in% c("CR", "PR", "SD") ~ AVALC.CTRSP,

      # PET-CT not done, not determined, or missing; use CT response if available
      (AVALC.PETRSP == "ND" | is.na(AVALC.PETRSP)) &
        AVALC.CTRSP %in% c("CR", "PR", "SD", "PD", "NED") ~ AVALC.CTRSP,

      # PET-CT not done or missing and CT is NE
      (AVALC.PETRSP == "ND" | is.na(AVALC.PETRSP)) & AVALC.CTRSP == "NE" ~ "NE",

      # No valid response available
      TRUE ~ "ND"
    ),
    PARAMCD = "OVRLRESC",
    PARAM = "Overall Response - Derived",
    PARAMN = 3,
    PARCAT1 = "LUGANO 2014"
  )
)
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adrs %>%
    filter(PARAMCD %in% c("OVRLRESC")) %>%
    arrange(!!!get_admiral_option("subject_keys"), AVISITN, PARAMN),
  display_vars = exprs(USUBJID, PARAMCD, PARAM, PARCAT1, AVALC, AVISIT, ADT)
)
```

### Derive `AVAL` (Numeric tumor response from `AVALC` values)

The `AVAL` values are not considered in the further parameter derivations below, and so changing `AVAL` here would not change the result of those derivations.

```{r}
adrs <- adrs %>%
  mutate(
    AVAL = case_when(
      AVALC == "CR" ~ 1,
      AVALC == "PR" ~ 2,
      AVALC == "SD" ~ 3,
      AVALC == "PD" ~ 4,
      AVALC == "NE" ~ 5,
      AVALC == "NED" ~ 6,
      AVALC == "PSP" ~ 7,
      AVALC == "ND" ~ 8,
      TRUE ~ NA_real_
    )
  )
```

```{r, echo=FALSE}
dataset_vignette(
  adrs %>%
    filter(PARAMCD %in% c("OVRLRESC")) %>%
    arrange(!!!get_admiral_option("subject_keys"), AVISITN, PARAMN),
  display_vars = exprs(USUBJID, PARAMCD, PARAM, PARCAT1, AVALC, AVAL, AVISIT, ADT)
)
```

## Other Endpoints {#other}

For examples of additional endpoints, such as Best Overall Response (BOR), Confirmed Best Overall Response (CBOR), and other oncology response endpoints, please see [Creating ADRS (Including Non-standard Endpoints)](adrs.html).
