---
title: "Creating ADRS"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating ADRS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This article describes creating an `ADRS` ADaM with common oncology endpoint parameters
based on RECIST v1.1. Therefore response values are expected as either `CR`, `PR`, `SD`,
`NON-CR/NON-PD`, `PD` or `NE`.

Examples are currently presented and tested using `ADSL` (ADaM) and `RS`, `TU` (SDTM) inputs.
However, other domains could be used. The functions and workflow could similarly be used
to create an intermediary `ADEVENT` ADaM.

**Note**: *All examples assume CDISC SDTM and/or ADaM format as input unless
otherwise specified.*

# Programming Workflow

* [Read in Data](#readdata)
* [Pre-processing of Input Records](#input)
* [Derive Progressive Disease Parameter](#pd)
* [Derive Response Parameter](#rsp)
* [Derive Clinical Benefit Parameter](#cb)
* [Derive Best Overall Response Parameter](#bor)
* [Derive Best Overall Response of CR/PR Parameter](#bcp)
* [Derive Response Parameters requiring Confirmation](#confirm)
* [Derive Parameters using Independent Review Facility (IRF) responses](#irf)
* [Derive Death Parameter](#death)
* [Derive Last Disease Assessment Parameters](#lsta)
* [Derive Measurable Disease at Baseline Parameter](#mdis)
* [Assign `ASEQ`](#aseq)
* [Add ADSL variables](#adsl_vars)

## Read in Data {#readdata}

To start, all data frames needed for the creation of `ADRS` should be read into
the environment.  This will be a company specific process.  Some of the 
data frames needed may be `ADSL`, `RS` and `TU`.

For example purpose, the SDTM and ADaM datasets (based on CDISC Pilot test data)---which are
included in `{admiral.test}`---are used.

```{r message=FALSE}
library(admiral)
library(admiralonco)
library(dplyr)
library(admiral.test)
library(lubridate)
library(stringr)
data("admiral_adsl")
data("admiral_rs")
data("admiral_tu")

adsl <- admiral_adsl
rs <- admiral_rs
tu <- admiral_tu

rs <- convert_blanks_to_na(rs)
tu <- convert_blanks_to_na(tu)
```
```{r echo=FALSE}
rs <- filter(rs, USUBJID %in% c("01-701-1015", "01-701-1023", "01-703-1086", "01-703-1096", "01-707-1037", "01-716-1024"))
tu <- filter(tu, USUBJID %in% c("01-701-1015", "01-701-1023", "01-703-1086", "01-703-1096", "01-707-1037", "01-716-1024"))
adsl <- filter(adsl, USUBJID %in% c("01-701-1015", "01-701-1023", "01-703-1086", "01-703-1096", "01-707-1037", "01-716-1024"))
```

At this step, it may be useful to join `ADSL` to your `RS` domain. Only the
`ADSL` variables used for derivations are selected at this step. The rest of the
relevant `ADSL` would be added later.

```{r eval=TRUE}
adsl_vars <- vars(RANDDT)
adrs <- derive_vars_merged(
  rs,
  dataset_add = adsl,
  new_vars = adsl_vars,
  by_vars = vars(STUDYID, USUBJID)
)
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, RSTESTCD, RSDTC, VISIT, RANDDT),
  filter = RSTESTCD == "OVRLRESP"
)
```

## Pre-processing of Input Records {#input}

The first step involves company-specific pre-processing of records for the required input to the downstream parameter functions.
Note that this could be needed multiple times (e.g. once for investigator and once for Independent Review Facility (IRF) records).
It could even involve merging input data from other sources besides `RS`, such as `ADTR`.

This step would include any required selection/derivation of `ADT` or applying any necessary partial date imputations,
updating `AVAL` (e.g. this should be ordered from best to worst response), and setting analysis flag `ANL01FL`. Common options for
`ANL01FL` would be to set null for invalid assessments or those occurring after new anti-cancer therapy, or to only flag assessments
on or after after date of first treatment/randomization, or rules to cover the case when a patient has multiple observations per visit
(e.g. by selecting worst value).
Another consideration could be extra potential protocol-specific sources of Progressive Disease such as radiological assessments,
which could be pre-processed here to create a PD record to feed downstream derivations.

The below shows an example of a possible company-specific implementation of this step.

### Select Overall Response Records and Set Parameter Details

In this case we use the overall response records from `RS` from the investigator as our starting point. The parameter details such
as `PARAMCD`, `PARAM` etc will always be company-specific, but an example is shown below so that you can trace through how these
records feed into the other parameter derivations.

```{r}
adrs <- adrs %>%
  filter(RSEVAL == "INVESTIGATOR" & RSTESTCD == "OVRLRESP") %>%
  mutate(
    PARAMCD = "OVR",
    PARAM = "Overall Response by Investigator",
    PARCAT1 = "Tumor Response",
    PARCAT2 = "Investigator",
    PARCAT3 = "Recist 1.1"
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, VISIT, RSTESTCD, RSEVAL, PARAMCD, PARAM, PARCAT1, PARCAT2, PARCAT3)
)
```

### Partial Date Imputation and Deriving `ADT`, `ADTF`, `AVISIT` etc

If your data collection allows for partial dates, you could apply a company-specific imputation rule at this stage
when deriving `ADT`. For this example, here we impute missing day to last possible date.

```{r}
adrs <- adrs %>%
  derive_vars_dt(
    dtc = RSDTC,
    new_vars_prefix = "A",
    highest_imputation = "D",
    date_imputation = "last"
  ) %>%
  mutate(AVISIT = VISIT)
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, RSSTRESC, RSDTC, ADT, ADTF)
)
```

### Derive `AVALC` and `AVAL`

Here we populate `AVALC` and create the numeric version as `AVAL` (ordered from best to worst response).
This ordering is already covered within our RECIST v1.1 parameter derivation functions, and so changing
`AVAL` here would not change the result of those derivations.

```{r}
adrs <- adrs %>%
  mutate(
    AVALC = RSSTRESC,
    AVAL = aval_resp(AVALC)
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, RSSTRESC, AVALC, AVAL)
)
```

### Derive `ANL01FL`

When deriving `ANL01FL` this is an opportunity to exclude any records that should not contribute to any downstream
parameter derivations. In the below example this includes only selecting valid assessments and those occurring on or
after randomization date.

```{r}
adrs <- adrs %>%
  mutate(
    ANL01FL = case_when(
      !is.na(AVAL) & ADT >= RANDDT ~ "Y",
      TRUE ~ NA_character_
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, RANDDT, ANL01FL)
)
```

Here is an alternative example where those records occurring after new anti-cancer therapy are additionally
excluded (where `NACTDT` would be pre-derived as first date of new anti-cancer therapy).

```{r, eval=FALSE}
adrs <- adrs %>%
  mutate(
    ANL01FL = case_when(
      !is.na(AVAL) & ADT >= RANDDT & ADT < NACTDT ~ "Y",
      TRUE ~ NA_character_
    )
  )
```

Note here that we don't filter out records after first `PD` at this stage, as that is specifically catered for in
the `{admiralonco}` parameter derivation functions in the below steps, via `source_pd` arguments.
However, if you prefer not to rely on `source_pd` arguments, then the user is free to filter out records after
first `PD` at this stage in a similar way via a `ANLzzFL` flag, and then you could leave `source_pd` as null in
all downstream parameter derivation function calls.

## Derive Progressive Disease Parameter {#pd}

Now that we have the input records prepared above with any company-specific requirements, we can start to derive
new parameter records. For the parameter derivations, all values except those overwritten by `set_values_to`
argument are kept from the earliest occurring input record fulfilling the required criteria.

The function `admiral::derive_param_first_event()` can be used to find the date of first `PD`.

```{r}
adrs <- adrs %>%
  derive_param_first_event(
    dataset_adsl = adsl,
    dataset_source = adrs,
    filter_source = PARAMCD == "OVR" & AVALC == "PD" & ANL01FL == "Y",
    date_var = ADT,
    set_values_to = vars(
      PARAMCD = "PD",
      PARAM = "Disease Progression by Investigator",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, ANL01FL),
  filter = PARAMCD == "PD"
)
```

For progressive disease, response and death parameters shown in steps here and below, in our
examples we show these as `ADRS` parameters, but they could equally be achieved via `ADSL` dates or `ADEVENT`
parameters.
If you prefer to store as an ADSL date, then the function `admiral::derive_var_extreme_dt()` could be used
to find the date of first `PD` as a variable, rather than as a new parameter record.
All the parameter derivation functions that use these dates are flexible to allow sourcing these
from any input source using `admiral::date_source()`. See examples below.

## Derive Response Parameter {#rsp}

The next required step is to define the source location for this newly derived `PD` date.

```{r}
pd <- date_source(
  dataset_name = "adrs",
  date = ADT,
  filter = PARAMCD == "PD" & AVALC == "Y"
)
```

An equivalent example if using `ADSL` instead could be as follows (where `PDDT` would be
pre-derived as first date of progressive disease).

```{r, eval=FALSE}
pd <- date_source(
  dataset_name = "adsl",
  date = PDDT
)
```

The function `derive_param_response()` can then be used to find the date of first response. This differs from
the `admiral::derive_param_first_event()` function in that it only looks for events occurring prior to first `PD`.
In the below example, the response condition has been defined as `CR` or `PR`.

```{r}
adrs <- adrs %>%
  derive_param_response(
    dataset_adsl = adsl,
    filter_source = PARAMCD == "OVR" & AVALC %in% c("CR", "PR") & ANL01FL == "Y",
    source_pd = pd,
    source_datasets = list(adrs = adrs),
    set_values_to = vars(
      PARAMCD = "RSP",
      PARAM = "Response by Investigator (confirmation not required)",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, ANL01FL),
  filter = PARAMCD == "RSP"
)
```

## Derive Clinical Benefit Parameter {#cb}

Similarly, we now define the source location for this newly derived first response date.

```{r}
resp <- date_source(
  dataset_name = "adrs",
  date = ADT,
  filter = PARAMCD == "RSP" & AVALC == "Y"
)
```

The function `derive_param_clinbenefit()` can then be used to derive the clinical benefit parameter, which we define
as a patient having had a response or a sustained period of time before first `PD`. This could also be known as
disease control. In this example the "sustained period" has been defined as 42 days after randomization date, using the
`ref_start_window` argument.

```{r}
adrs <- adrs %>%
  derive_param_clinbenefit(
    dataset_adsl = adsl,
    filter_source = PARAMCD == "OVR" & ANL01FL == "Y",
    source_resp = resp,
    source_pd = pd,
    source_datasets = list(adrs = adrs),
    reference_date = RANDDT,
    ref_start_window = 42,
    set_values_to = vars(
      PARAMCD = "CB",
      PARAM = "Clinical Benefit by Investigator (confirmation for response not required)",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, RANDDT, ANL01FL),
  filter = PARAMCD == "CB"
)
```

## Derive Best Overall Response Parameter {#bor}

The function `derive_param_bor()` can be used to derive the best overall response (without confirmation required) parameter.
Similar to the above function you can optionally decide what period would you consider a `SD` or `NON-CR/NON-PD` as being
eligible from. In this example, 42 days after randomization date has been used again.

```{r}
adrs <- adrs %>%
  derive_param_bor(
    dataset_adsl = adsl,
    filter_source = PARAMCD == "OVR" & ANL01FL == "Y",
    source_pd = pd,
    source_datasets = list(adrs = adrs),
    reference_date = RANDDT,
    ref_start_window = 42,
    set_values_to = vars(
      PARAMCD = "BOR",
      PARAM = "Best Overall Response by Investigator (confirmation not required)",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, RANDDT, ANL01FL),
  filter = PARAMCD == "BOR"
)
```

Note that the above gives pre-defined `AVAL` values of:
`"CR" ~ 1`, `"PR" ~ 2`, `"SD" ~ 3`, `"NON-CR/NON-PD" ~ 4`, `"PD" ~ 5`, `"NE" ~ 6`, `"MISSING" ~ 7`.

If you'd like to provide your own company-specific ordering here you could do this as follows:

```{r, eval=FALSE}
aval_resp_new <- function(arg) {
  assert_character_vector(arg)
  case_when(
    arg == "CR" ~ 7,
    arg == "PR" ~ 6,
    arg == "SD" ~ 5,
    arg == "NON-CR/NON-PD" ~ 4,
    arg == "PD" ~ 3,
    arg == "NE" ~ 2,
    arg == "MISSING" ~ 1,
    TRUE ~ NA_real_
  )
}
```

Then add the additional argument `aval_fun = aval_resp_new` to the above `derive_param_bor()` call.

## Derive Best Overall Response of CR/PR Parameter {#bcp}

The function `derive_param_first_event()` can be used to check if a patient had a response for BOR.

```{r}
adrs <- adrs %>%
  derive_param_first_event(
    dataset_adsl = adsl,
    dataset_source = adrs,
    filter_source = PARAMCD == "BOR" & AVALC %in% c("CR", "PR") & ANL01FL == "Y",
    date_var = ADT,
    set_values_to = vars(
      PARAMCD = "BCP",
      PARAM = "Best Overall Response of CR/PR by Investigator (confirmation not required)",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, ANL01FL),
  filter = PARAMCD == "BCP"
)
```

## Derive Response Parameters requiring Confirmation {#confirm}

Any of the above response parameters can be repeated for "confirmed" responses only. For these the functions
`derive_param_confirmed_resp()` and `derive_param_confirmed_bor()` can be used. Here is an example of a BOR parameter
when requiring confirmation. The assessment and the confirmatory assessment need to occur at least 28 days apart, using the
`ref_confirm` argument.

```{r}
adrs <- adrs %>%
  derive_param_confirmed_bor(
    dataset_adsl = adsl,
    filter_source = PARAMCD == "OVR" & ANL01FL == "Y",
    source_pd = pd,
    source_datasets = list(adrs = adrs),
    reference_date = RANDDT,
    ref_start_window = 42,
    ref_confirm = 28,
    set_values_to = vars(
      PARAMCD = "CBOR",
      PARAM = "Best Overall Response by Investigator",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, RANDDT, ANL01FL),
  filter = PARAMCD == "CBOR"
)
```

## Derive Parameters using Independent Review Facility (IRF) responses {#irf}

All of the above steps can be repeated for different sets of records, such as now using assessments from the Independent
Review Facility instead of investigator. For this you would just need to replace the first steps with selecting the required
records, and then feed these as input to the downstream parameter functions.

Remember that a new progressive disease and response source object would be required for passing to
`source_pd` and `source_resp` respectively.

```{r}
adrsirf <- rs %>%
  filter(RSEVAL == "INDEPENDENT ASSESSOR" & RSEVALID == "RADIOLOGIST 1" & RSTESTCD == "OVRLRESP") %>%
  mutate(
    PARAMCD = "OVRR1",
    PARAM = "Overall Response by Radiologist 1",
    PARCAT1 = "Tumor Response",
    PARCAT2 = "Radiologist",
    PARCAT3 = "Recist 1.1"
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrsirf,
  display_vars = vars(USUBJID, VISIT, RSTESTCD, RSEVAL, PARAMCD, PARAM, PARCAT1, PARCAT2, PARCAT3),
  filter = PARAMCD == "OVRR1"
)
```

Then in all the calls to the parameter derivation functions you would replace the `PARAMCD == "OVR"` source with
`PARAMCD == "OVRR1"`.

## Derive Death Parameter {#death}

The function `admiral::derive_param_first_event()` can be used to create a new death parameter using death date from `ADSL`.
We need to restrict the columns from `ADSL` as we'll merge all required variables later across all our `ADRS` records.

```{r}
adsldth <- adsl %>%
  select(STUDYID, USUBJID, DTHDT, !!!adsl_vars)

adrs <- adrs %>%
  derive_param_first_event(
    dataset_adsl = adsldth,
    dataset_source = adsldth,
    filter_source = !is.na(DTHDT),
    date_var = DTHDT,
    set_values_to = vars(
      PARAMCD = "DEATH",
      PARAM = "Death",
      PARCAT1 = "Reference Event",
      ANL01FL = "Y"
    )
  ) %>%
  select(-DTHDT)
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, ANL01FL),
  filter = PARAMCD == "DEATH"
)
```

## Derive Last Disease Assessment Parameters {#lsta}

The function `derive_param_lasta()` can be used to create a parameter for last disease assessment. There is flexibility
to use `source_pd` argument to only look up to first `PD` assessment, but in the below example all assessments are included.

```{r}
adrs <- adrs %>%
  derive_param_lasta(
    filter_source = PARAMCD == "OVR" & ANL01FL == "Y",
    set_values_to = vars(
      PARAMCD = "LSTA",
      PARAM = "Last Disease Assessment by Investigator",
      PARCAT1 = "Tumor Response",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, ANL01FL),
  filter = PARAMCD == "LSTA"
)
```

## Derive Measurable Disease at Baseline Parameter {#mdis}

The function `admiral::derive_param_exist_flag()` can be used to check whether a patient has measurable disease at baseline,
according to a company-specific condition. In this example we check `TU` for target lesions during the baseline visit.
We need to restrict the columns from `ADSL` as we'll merge all required variables later across all our `ADRS` records.

```{r}
adslmdis <- adsl %>%
  select(STUDYID, USUBJID, !!!adsl_vars)

adrs <- adrs %>%
  derive_param_exist_flag(
    dataset_adsl = adslmdis,
    dataset_add = tu,
    condition = TUEVAL == "INVESTIGATOR" & TUSTRESC == "TARGET" & VISIT == "BASELINE",
    false_value = "N",
    missing_value = "N",
    set_values_to = vars(
      PARAMCD = "MDIS",
      PARAM = "Measurable Disease at Baseline by Investigator",
      PARCAT2 = "Investigator",
      PARCAT3 = "Recist 1.1",
      ANL01FL = "Y"
    )
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, AVISIT, PARAMCD, PARAM, AVALC, ADT, ANL01FL),
  filter = PARAMCD == "MDIS"
)
```

## Assign `ASEQ` {#aseq}

The function `admiral::derive_var_obs_number()` can be used to derive `ASEQ`. An example call is:

```{r eval=TRUE}
adrs <- adrs %>%
  derive_var_obs_number(
    by_vars = vars(STUDYID, USUBJID),
    order = vars(PARAMCD, ADT, VISITNUM, RSSEQ),
    check_type = "error"
  )
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, PARAMCD, ADT, VISITNUM, AVISIT, ASEQ),
  filter = USUBJID == "01-701-1015"
)
```

## Add ADSL variables {#adsl_vars}

If needed, the other `ADSL` variables can now be added.
List of ADSL variables already merged held in vector `adsl_vars`.

```{r eval=TRUE}
adrs <- adrs %>%
  derive_vars_merged(
    dataset_add = select(adsl, !!!negate_vars(adsl_vars)),
    by_vars = vars(STUDYID, USUBJID)
  )
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = vars(USUBJID, RFSTDTC, RFENDTC, DTHDTC, DTHFL, AGE, AGEU),
  filter = USUBJID == "01-701-1015"
)
```

# Example Script {#example}

ADaM | Sample Code
---- | --------------
ADRS | [ad_adrs.R](https://github.com/pharmaverse/admiralonco/blob/test/inst/templates/ad_adrs.R){target="_blank"}
