---
title: "Creating ADTTE"
output:
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Creating ADTTE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This article describes creating an `ADTTE` (time-to-event) ADaM with common oncology
endpoint parameters.

The main part in programming a time-to-event dataset is the definition of the
events and censoring times. `{admiral}`/`{admiralonco}` supports single events like
death (Overall Survival) or composite events like disease progression or death (Progression
Free Survival). More than one source dataset can be used for the definition of the event
and censoring times.

The majority of the functions used here exist from `{admiral}`, except for the
`tte_sources` helper object, provided as an example from `{admiralonco}`.
In practice, each company would create their own version of this, as likely the
exact specifications such as filtering condition or description metadata will vary.

**Note**: *All examples assume CDISC SDTM and/or ADaM format as input unless
otherwise specified.*

## Required Packages

The examples of this vignette require the following packages.

```{r, warning=FALSE, message=FALSE}
library(admiral)
library(admiralonco)
library(dplyr)
library(admiral.test)
```

```{r, warning=FALSE, message=FALSE, include=FALSE}
library(lubridate)
```

# Programming Workflow

* [Read in Data](#readdata)
* [Derive Parameters (`CNSR`, `ADT`, `STARTDT`)](#parameters)
* [Derive Analysis Value (`AVAL`)](#aval)
* [Derive Analysis Sequence Number (`ASEQ`)](#aseq)
* [Add ADSL Variables](#adslvars)

## Read in Data {#readdata}

To start, all datasets needed for the creation of the time-to-event dataset
should be read into the environment. This will be a company specific process.

For example purpose, the ADaM datasets---which are included
in `{admiral}` and `{admiralonco}`---are used. An alternative might be to use
`ADEVENT` as input.

```{r}
data("admiral_adsl")
data("admiral_adrs")
adsl <- admiral_adsl
adrs <- admiral_adrs
```

## Derive Parameters (`CNSR`, `ADT`, `STARTDT`) {#parameters}

To derive the parameter dependent variables like `CNSR`, `ADT`, `STARTDT`,
`EVNTDESC`, `SRCDOM`, `PARAMCD`, ... the `admiral::derive_param_tte()` function can be
used. It adds one parameter to the input dataset with one observation per
subject. Usually it is called several times.

For each subject it is determined if an event occurred. In the affirmative the
analysis date `ADT` is set to the earliest event date. If no event occurred, the
analysis date is set to the latest censoring date.

The events and censorings are defined by the `admiral::event_source()` and the
`admiral::censor_source()` class respectively. It defines

- which observations (`filter` parameter) of a source dataset (`dataset_name`
parameter) are potential events or censorings,
- the value of the `CNSR` variable (`censor` parameter), and
- which variable provides the date (`date` parameter).

The date can be provided as date (`--DT` variable), datetime (`--DTM` variable),
or character ISO-8601 date (`--DTC` variable).

CDISC strongly recommends `CNSR = 0` for events and positive integers for
censorings. `{admiral}`/`{admiralonco}` enforce this recommendation. Therefore the `censor`
parameter is available for `admiral::censor_source()` only. It is defaulted to `1`.

The `dataset_name` parameter expects a character value which is used as an
identifier. The actual data which is used for the derivation of the parameter is
provided via the `source_datasets` parameter of `admiral::derive_param_tte()`. It expects
a named list of datasets. The names correspond to the identifiers specified for
the `dataset_name` parameter. This allows to define events and censoring
independent of the data.

### Pre-Defined Time-to-Event Source Objects

The table below shows all pre-defined `tte_source` objects which should cover the most
common oncology use cases.

```{r echo=FALSE}
knitr::kable(list_tte_source_objects(package = "admiralonco"))
```

As mentioned in the introduction, each company would create their own version of this with
the required filtering conditions and metadata as per your company approach. An example of
a possible different approach could be as follows, where death is sourced from `ADSL`,
instead of `ADRS`, and the given `EVNTDESC` is different.

```{r, eval=FALSE}
adsl_death_event <- event_source(
  dataset_name = "adsl",
  date = DTHDT,
  set_values_to = vars(
    EVNTDESC = "STUDY DEATH",
    SRCDOM = "ADSL",
    SRCVAR = "DTHDT"
  )
)
```

An optional step at this stage would be required to enable derivation of duration of
response: If using `ADRS` / `ADEVENT` parameters as input for any response dates
(instead of a variable in `ADSL`) then you would need to use `admiral::derive_vars_merged()`
to add the response date as a temporary variable (e.g. `TEMP_RESPDT`) to be able to feed
into `admiral::derive_param_tte()` as the start date. You would also need to use this
to filter the input datasets so as to only derive the records for responders.
This could also be repeated as needed for IRF and confirmed responses.

Here is an example of the code needed.

```{r}
adsl <- adsl %>%
  derive_vars_merged(
    dataset_add = adrs,
    filter_add = PARAMCD == "RSP" & AVALC == "Y" & ANL01FL == "Y",
    by_vars = vars(STUDYID, USUBJID),
    new_vars = vars(TEMP_RESPDT = ADT)
  )

adrs <- adrs %>%
  derive_vars_merged(
    dataset_add = adsl,
    by_vars = vars(STUDYID, USUBJID),
    new_vars = vars(TEMP_RESPDT)
  )
```

The pre-defined objects can be passed directly to `admiral::derive_param_tte()` to create
a new time-to-event parameter. Below shows example calls for Overall Survival (OS), Progression
Free Survival (PFS), and duration of response (as above, this is only derived for responder
patients so we have to filter source datasets). Note that the reason for including a
randomization date censor is to catch those patients that never have a tumor assessment.

```{r}
adtte <- derive_param_tte(
  dataset_adsl = adsl,
  start_date = RANDDT,
  event_conditions = list(death_event),
  censor_conditions = list(lastalive_censor, rand_censor),
  source_datasets = list(adsl = adsl, adrs = adrs),
  set_values_to = vars(PARAMCD = "OS", PARAM = "Overall Survival")
) %>%
  derive_param_tte(
    dataset_adsl = adsl,
    start_date = RANDDT,
    event_conditions = list(pd_event, death_event),
    censor_conditions = list(lasta_censor, rand_censor),
    source_datasets = list(adsl = adsl, adrs = adrs),
    set_values_to = vars(PARAMCD = "PFS", PARAM = "Progression Free Survival")
  ) %>%
  derive_param_tte(
    dataset_adsl = adsl,
    start_date = TEMP_RESPDT,
    event_conditions = list(pd_event, death_event),
    censor_conditions = list(lasta_censor),
    source_datasets = list(adsl = filter(adsl, !is.na(TEMP_RESPDT)), adrs = filter(adrs, !is.na(TEMP_RESPDT))),
    set_values_to = vars(PARAMCD = "RSD", PARAM = "Duration of Response")
  )
```
```{r, echo=FALSE}
dataset_vignette(
  adtte,
  display_vars = vars(USUBJID, PARAMCD, PARAM, STARTDT, ADT, CNSR)
)
```

### Creating Your Own Time-to-Event Source Objects

We advise you consult the `{admiral}`
["Creating a BDS Time-to-Event ADaM" vignette](https://pharmaverse.github.io/admiral/articles/bds_tte.html)
for further guidance on the different options available and more examples.

One extra common oncology case we include here is around PFS when censoring
at next anti-cancer therapy. This could either be controlled using `ANLzzFL` as
explained in the ADRS vignette, so that records after next anti-cancer therapy
never contribute to the PD and DEATH parameters. Or alternatively you can control
this on the ADTTE side by filtering which records are used in
`admiral::event_source()` and `admiral::censor_source()`.
The censor could be set as whichever date your analysis requires,
e.g. date of last tumor assessment prior to next anti-cancer therapy or last
radiological assessment.
This case is demonstrated in the below example (where `NACTDT` would be
pre-derived as first date of next anti-cancer therapy and `LASTANDT` as date of
last assessment prior to next anti-cancer therapy).

```{r, eval=FALSE}
pd_nact_event <- event_source(
  dataset_name = "adsl",
  filter = PDDT < NACTDT,
  date = PDDT,
  set_values_to = vars(
    EVNTDESC = "Disease Progression prior to NACT",
    SRCDOM = "ADSL",
    SRCVAR = "PDDT"
  )
)

death_nact_event <- event_source(
  dataset_name = "adsl",
  filter = DTHDT < NACTDT,
  date = DTHDT,
  set_values_to = vars(
    EVNTDESC = "Death prior to NACT",
    SRCDOM = "ADSL",
    SRCVAR = "DTHDT"
  )
)

lasta_nact_censor <- censor_source(
  dataset_name = "adsl",
  date = LASTANDT,
  set_values_to = vars(
    EVNTDESC = "Last Tumor Assessment prior to NACT",
    CNSDTDSC = "Last Tumor Assessment prior to NACT",
    SRCDOM = "ADSL",
    SRCVAR = "LASTANDT"
  )
)

adtte <- derive_param_tte(
  dataset_adsl = adsl,
  start_date = RANDDT,
  event_conditions = list(pd_nact_event, death_nact_event),
  censor_conditions = list(lasta_nact_censor, rand_censor),
  source_datasets = list(adsl = adsl),
  set_values_to = vars(PARAMCD = "PFSNACT", PARAM = "Progression Free Survival prior to NACT")
)
```

## Derive Analysis Value (`AVAL`) {#aval}

The analysis value (`AVAL`) can be derived by calling `admiral::derive_vars_duration()`.

This example derives the time to event in days. Other units can be requested by
specifying the `out_unit` parameter.

```{r eval=TRUE}
adtte <- adtte %>%
  derive_vars_duration(
    new_var = AVAL,
    start_date = STARTDT,
    end_date = ADT
  )
```

```{r, echo=FALSE}
dataset_vignette(
  adtte
)
```

## Derive Analysis Sequence Number (`ASEQ`) {#aseq}

The `{admiral}` function `admiral::derive_var_obs_number()` can be used to derive `ASEQ`:

```{r eval=TRUE}
adtte <- adtte %>%
  derive_var_obs_number(
    by_vars = vars(STUDYID, USUBJID),
    order = vars(PARAMCD),
    check_type = "error"
  )
```

```{r, echo=FALSE}
dataset_vignette(adtte)
```

## Add ADSL Variables {#adslvars}

Variables from ADSL which are required for time-to-event analyses, e.g.,
treatment variables or covariates can be added using `admiral::derive_vars_merged()`.

```{r eval=TRUE}
adtte <- adtte %>%
  derive_vars_merged(
    dataset_add = adsl,
    new_vars = vars(ARMCD, ARM, ACTARMCD, ACTARM, AGE, SEX),
    by_vars = vars(STUDYID, USUBJID)
  )
```

```{r, echo=FALSE}
dataset_vignette(
  adtte,
  display_vars = vars(USUBJID, PARAMCD, CNSR, AVAL, ARMCD, AGE, SEX)
)
```

# Example Script {#example}

ADaM | Sample Code
---- | --------------
ADTTE | [ad_adtte.R](https://github.com/pharmaverse/admiralonco/blob/test/inst/templates/ad_adtte.R){target="_blank"}
