---
title: "Creating and Using New Anti-Cancer Start Date"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating and Using New Anti-Cancer Start Date}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(admiralonco)
link <- function(text, url) {
  return(
    paste0(
      "[", text, "]",
      "(", url, ")"
    )
  )
}
dyn_link <- function(text,
                     base_url,
                     relative_url = "",
                     # Change to TRUE when admiral adopts multiversion docs
                     is_multiversion = FALSE,
                     multiversion_default_ref = "main") {
  url <- paste(base_url, relative_url, sep = "/")
  if (is_multiversion) {
    url <- paste(
      base_url,
      Sys.getenv("BRANCH_NAME", multiversion_default_ref),
      relative_url,
      sep = "/"
    )
  }
  return(link(text, url))
}
# Other variables
admiral_homepage <- "https://pharmaverse.github.io/admiral/cran-release"

library(admiraldev)
```

# Introduction

In Oncology trials, for censoring the patients for any time-to-event analysis (e.g., progression free survival analysis, duration of response, etc.), it is necessary to check if the patients have started any new anti-cancer therapy. Anti-cancer therapy generally includes radiation therapy, cancer related surgery and any systemic anti-cancer therapy such as chemotherapy, t-cell therapy, etc.
These therapies can be collected on a prior or on-treatment CRF pages, with radiation and surgery being mapped to the SDTM PR domain and systemic anti-cancer therapy being mapped to the SDTM CM domain. Collection and mapping may vary with each company and it is important that only on-treatment therapies are considered for the censoring of time-to-event. Generally, if the patients start a new anti-cancer therapy that were on-treatment, they should be discontinued from the study treatment and no further efficacy assessments should be performed.

This article describes a process for creating a new anti-cancer start date from a single and multiple source SDTMs.

# Programming Workflow

-   [Read in Data and create test data](#readdata)
-   [Single Source](#input)
-   [Multiple Sources - Prerequisite Step](#multiinput)
-   [Multiple Sources - Derive New Anti-Cancer Start Date and Merge with `ADSL`](#nactdt)
-   [Using the New Anti-Cancer Therapy Date](#unactdt)
-   [Derive Date Parameter](#par)

## Read in Data {#readdata}

To start, all data frames needed for the creation of the new anti-cancer therapy start date (`NACTDT`) should be read into the environment. This will be a company specific process. Some of the data frames needed may be `CM` and `PR`.

For example purpose, `CM` and `PR` has been generated (based on CDISC Pilot test data), with oncology specific test data.

```{r, warning=FALSE, message=FALSE}
library(admiral)
library(dplyr)
library(admiral.test)

adsl <- admiral_adsl

cm <- tribble(
  ~STUDYID, ~USUBJID, ~CMCAT, ~CMSCAT, ~CMTRT, ~CMSTDTC,
  "CDISCPILOT01", "01-701-1015", "PRIOR TREATMENT", "CHEMOTHERAPY", "DEXRAZOXANE", NA,
  "CDISCPILOT01", "01-701-1015", "ON TREATMENT", "CHEMOTHERAPY", "DEXROZOXANE", "2014-07-02",
  "CDISCPILOT01", "01-701-1015", "ON TREATMENT", "CHEMOTHERAPY", "DEXROZOXANE", "2014-06-19",
  "CDISCPILOT01", "01-701-1028", "PRIOR TREATMENT", "CHEMOTHERAPY", "METHOTREXATE", NA,
  "CDISCPILOT01", "01-701-1028", "ON TREATMENT", "CHEMOTHERAPY", "METHOTREXATE", "2014-01-14",
  "CDISCPILOT01", "01-701-1034", "PRIOR TREATMENT", "CHEMOTHERAPY", "OLAPARIB", NA,
  "CDISCPILOT01", "01-701-1034", "ON TREATMENT", "CHEMOTHERAPY", "OLAPARIB", "2014-12-30",
  "CDISCPILOT01", "01-701-1097", "PRIOR TREATMENT", "CHEMOTHERAPY", "TEMODAL", NA,
  "CDISCPILOT01", "01-701-1097", "ON TREATMENT", "CHEMOTHERAPY", "TEMODAL", "2013-12-31",
)

pr <- tribble(
  ~STUDYID, ~USUBJID, ~PRCAT, ~PRSCAT, ~PRTRT, ~PRSTDTC,
  "CDISCPILOT01", "01-701-1015", "CANCER RELATED", "ON TREATMENT", "SURGERY", "2014-06-18",
  "CDISCPILOT01", "01-701-1034", "CANCER RELATED", "ON TREATMENT", "SURGERY", "2014-12-16",
  "CDISCPILOT01", "01-701-1028", "CANCER RELATED", "PRIOR TREATMENT", "SURGERY", NA,
)
```

## Single Source{#input}

To derive the New Anti-Cancer Therapy start date with data from a single source, the function `admiral::derive_vars_merged()` can be used. If numeric date is required,  `admiral::convert_dtc_to_dt` , or if time part is needed, `admiral::convert_dtc_to_dtm()` can be used as part of an expression for the `new_vars` argument.

```{r message=FALSE}

adsl <- derive_vars_merged(
  admiral_adsl,
  dataset_add = cm,
  by_vars = exprs(STUDYID, USUBJID),
  order = exprs(NACTDT),
  mode = "first",
  new_vars = exprs(NACTDT = convert_dtc_to_dt(CMSTDTC)),
  filter_add = CMSCAT == "CHEMOTHERAPY" & CMCAT == "ON TREATMENT"
)
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adsl,
  display_vars = exprs(USUBJID, NACTDT),
  filter = !is.na(NACTDT)
)
```

## Multiple Sources - Prerequisite Steps {#multiinput}

To derive the New Anti-Cancer Therapy start date with data from multiple sources, the function `admiral::derive_var_extreme_dt()` can be used, or if time part is needed, `admiral::derive_var_extreme_dtm()` can be used.
 
 
### Prerequisite 

Prior to calling the function, an input `date_source` object must be created for each source dataset using `admiral::date_source()`. Within this function, any company-specific subsetting can be passed to the `filter` argument. If numeric version of the dates will be needed, `admiral::convert_dtc_to_dt()`, or if time part is needed, `admiral::convert_dtc_to_dtm()` can be used as part of an expression for the `new_vars` argument. 

```{r message=FALSE}
cm_date <- date_source(
  dataset_name = "cm",
  filter = CMSCAT == "CHEMOTHERAPY" & CMCAT == "ON TREATMENT" & !is.na(CMSTDTC),
  date = convert_dtc_to_dt(CMSTDTC)
)

pr_date <- date_source(
  dataset_name = "pr",
  filter = PRCAT == "CANCER RELATED" & PRSCAT == "ON TREATMENT" & !is.na(PRSTDTC),
  date = convert_dtc_to_dt(PRSTDTC)
)
```

## Multiple Sources - Derive New Anti-Cancer Start Date and Merge with `ADSL` {#nactdt}

After completion of the prerequisite step, the new anti-cancer date can be derived while simultaneously adding it to `ADSL`.

```{r message=FALSE}
adsl <- admiral_adsl %>%
  derive_var_extreme_dt(
    new_var = NACTDT,
    cm_date, pr_date,
    source_datasets = list(
      cm = cm,
      pr = pr
    ),
    mode = "first"
  )
```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adsl,
  display_vars = exprs(USUBJID, NACTDT),
  filter = !is.na(NACTDT)
)
```

## Using the New Anti-Cancer Therapy Date {#unactdt}

Please refer to `{admiralonco}` [Derive `ANL01FL`](adrs.html#anl01fl) for an example on the usage of `NACTDT`. Additionally, `{admiralonco}` [Creating Your Own Time-to-Event Source Objects](adtte.html#tteobj) may be referenced for case usage.

## Derive Date Parameter {#par}

As of `{admiral}` version 0.11.0, `admiral::derive_param_extreme_record()` is recommended to be used for deriving parameter based on the first or last observation from single and multiple sources. 

Based on individual company standards, this parameter may be added to an event or date dataset such as `ADEVENT` or `ADDATES`, which are generally intermediate datasets to `ADTTE`. For demonstration purpose, a new anti-cancer date parameter will be merged into `ADRS`. A list of `record_source()` objects is expected and this will contain the sources of the dates and any company specific filtering.

```{r, eval=TRUE, echo=TRUE}

library(admiralonco)

adrs <- derive_param_extreme_record(
  dataset = admiral_adrs,
  sources = list(
    records_source(
      dataset_name = "cm",
      filter = CMSCAT == "CHEMOTHERAPY" & CMCAT == "ON TREATMENT" & !is.na(CMSTDTC),
      new_vars = exprs(
        ADT = convert_dtc_to_dt(CMSTDTC),
        AVALC = CMTRT
      )
    ),
    records_source(
      dataset_name = "pr",
      filter = PRCAT == "CANCER RELATED" & PRSCAT == "ON TREATMENT" & !is.na(PRSTDTC),
      new_vars = exprs(
        ADT = convert_dtc_to_dt(PRSTDTC),
        AVALC = PRTRT
      )
    )
  ),
  source_datasets = list(cm = cm, pr = pr),
  by_vars = exprs(STUDYID, USUBJID),
  order = exprs(ADT),
  mode = "first",
  set_values_to = exprs(
    PARAMCD = "NACTDT",
    PARAM = "New Anti-Cancer Therapy Start Date"
  )
)

```

```{r, eval=TRUE, echo=FALSE}
dataset_vignette(
  adrs,
  display_vars = exprs(USUBJID, PARAMCD, PARAM, ADT, AVALC),
  filter = !is.na(ADT) & PARAMCD == "NACTDT"
)
```
